National Pipeline System as a Solution to Future Water Shortage in the United States

### 2013-02-04 paper.tex

Our model projects levels of supply and demand in 2025, and proceeds
to recommend infrastructure to meet that demand while minimising cost.
This reflects the main precedent in major development of water supply
infrastructure in the United States, namely the \$24.6 billion Clean
Water Act \cite{key-15}, established with the aim of eliminating
all pollution of water supplies.

@@

There are two anticipated causes of significant variance in water
levels:
\begin{itemize}
\item Pathological rainfall levels (ie. extended periods of drought) 
\item Significant unexpected change in population or behaviour 
\end{itemize}
Water systems in general are resistant to small variance in water
supply or demand, and our system in particular fares well in this
regard. Many of our suggested developments are particularly well equipped
to handle over-supply: our model recommends the construction of additional
pipelines and de-salinization plants — pipeline flow can be easily
reduced, while de-salinization plants can be set to operate at partial
capacity, saving electricity and avoiding over-supply.

@@

While de-salinization is simply not an option at all for much of central
America, one idea we considered was the construction of additional
dams and reservoirs. One considerable environmental issue associated
with large dams is that they can cause unanticipated changes in overall
river flow\cite{key-17}, totalling potentially huge amounts in damages.

A secondary and extremely desirable effect of high national connectivity is decreased reliance on groundwater. As a case study to understand the overexploitation of aquifers, consider the High Plains aquifer, which supplies water to Colorado, Kansas, Nebraska, New Mexico, Oklahoma, South Dakota, Texas, and Wyoming, some of the most agricultural states in the US \cite{key-19}.
Due to overexploitation many areas have experienced severe water-level decline, an average of 12.8 feet. While 9 percent of areas suffered decline as severe as 50 feet, only a bare 2 percent saw any significant (over 10 feet) of water-level rise since predevelopment. It is estimated that 9 percent of the total storage in this massive and essential groundwater supply was exhausted by 2005.

>>>>

Our model projects levels of supply and demand in 2025, and proceeds
to recommend infrastructure to meet that demand while minimizing cost.
This reflects the main precedent in major development of water supply
infrastructure in the United States, namely the \$24.6 billion Clean
Water Act \cite{key-15}, established with the aim of eliminating
all pollution of water supplies.

@@

There are two anticipated causes of significant variance in water
levels:
\begin{itemize}
\item Pathological rainfall levels (ie. extended periods of drought) 
\item Significant unexpected change in population or behavior 
\end{itemize}
Water systems in general are resistant to small variance in water
supply or demand, and our system in particular fares well in this
regard. Many of our suggested developments are particularly well equipped
to handle over-supply: our model recommends the construction of additional
pipelines and de-salinization plants — pipeline flow can be easily
reduced, while de-salinization plants can be set to operate at partial
capacity, saving electricity and avoiding over-supply.

@@

While de-salinization is simply not an option at all for much of central
America, one idea we considered was the construction of additional
dams and reservoirs. One considerable environmental issue associated
with large dams is that they can cause unanticipated changes in overall
river flow\cite{key-17}, totaling potentially huge amounts in damages.

A secondary and extremely desirable effect of high national connectivity is decreased reliance on groundwater. As a case study to understand the overexploitation of aquifers, consider the High Plains aquifer, which supplies water to Colorado, Kansas, Nebraska, New Mexico, Oklahoma, South Dakota, Texas, and Wyoming, some of the most agricultural states in the US \cite{key-19}.
Due to overexploitation many areas have experienced severe water-level decline, an average of 12.8 feet. While 9 percent of areas suffered decline as severe as 50 feet, only a bare 2 percent saw any significant (over 10 feet) degree of water-level rise since  prior to development. It is estimated that 9 percent of the total storage in this massive and essential groundwater supply was exhausted in 2005.
